---
title: "Interacting with the UNFCCC API using remis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interacting with the UNFCCC API using remis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`remis` uses the UNFCCC Data Interface Flexible Query API to download Inventory Data.
At a later stage, functions for downloading Compilation and Accounting Data (`cad` and `cadCP2`) will be included.  
Data downloads via queries require a variable for a specific party and year, which are referred to via unique `id`'s.
While party and year `id`'s are linked to respective data in a fairly straight-forward manner (e.g., single country or year), variable ids are unique descriptors for a combination of data sources consisting of:

- category
- classification
- measurement
- gas
- unit

The above are referred to as *ccmgu*'s.

Initializing `remis` with `rem_init` provides a list with several objects to inspect available data and query for it.
These include `id` overviews for parties, years, *ccmgu*'s, as well as respective 
variable tables and the query object for requests `.req`.


```{r setup, meassage = FALSE}
library(remis)

rem <- rem_init()

names(rem)
```

Note that only calls using the Flexible Query API are currently implemented. 
These allow interacting with data sources from:

```{r results='asis', echo=FALSE}

cat(paste0('- ', names(rem$categories[1:3]), collapse = '\n'))

```

Queries for data within `r paste(names(rem$categories[4:5]), collapse = ' and ')` will be available in future versions.


Further details on the objects within `rem` are:

```{r results='asis', echo=FALSE}
cat(
  paste0('- **',
         names(rem),
         '** - class: ',
         unlist(sapply(rem, function(x) paste(class(x), collapse = ', '))),
         collapse = '\n')
)

```



## Building up Flexible Queries

Details on the Flexible Query API can be found in the official documentation of the [Data Interface for UNFCCC](https://unfccc.int/process-and-meetings/transparency-and-reporting/greenhouse-gas-data/data-interface-help#Flexible-Queries-).    

Define whether annexOne or nonAnnexOne data shall be downloaded, and then:

1. Select party
2. Select year
3. Choose variable based on `ccmgu`'s
4. Execute query

Note, that objects in `rem` (excluding `parties`) provide all data sources (where applicable) reflected in the list items `r names(rem$categories)`.

### Example

#### **1. Select Party**

Here we choose Germany and France.

```{r example-party}

# in RStudio: 
# View(rem$parties)

parties_mask <- rem$parties$parties_code %in% c('DEU', 'FRA')
parties_id <- rem$parties[parties_mask, c('id', 'name', 'categoryCode')]
knitr::kable(parties_id)

```
Note, that both parties are in Annex I.

#### **2. Select years**

We will select all years from 1990 to the latest inventory year (2020).
```{r example-years}
# base year has id = 0
years <- rem$years$annexOne[rem$years$annexOne$id != 0, ]
knitr::kable(years)


```

#### **3. Select from ccmug's**

The selection of *ccmug*'s is possible is several ways.
Ideally, one item of interest should be selected, such as a category or a measure, through which the appropriate variable for querying is identified.
This can be done either by browsing in `RStudio` with `View(rem)` and using the interactive filter function, or by searching programmatically. 
The example below shows the latter.

Here, we assume an interest in missions from cars in the transportation sector by fuel type and in total in the CRF-Category *1.A.3.b.i  Cars* (id = 9279).
Therefore, we begin with *category* as our starting point.


```{r example-ccmug}

# parties are annexOne, thus:

category_mask <- grepl('1.A.3.b.i  Cars', rem$categories$annexOne$name)
category <- rem$categories$annexOne[category_mask, ]
knitr::kable(as.data.frame(t(category)))

```
The table above also highlights the nesting structure throughout the CRF-Category tree.

Next, find appropriate variables (i.e. combination of *ccmug*'s_ that contain our category of interest.

```{r example-variables}
variables <- select_varid(
  vars = rem$variables$annexOne,
  category_id = category$id)
knitr::kable(head(variables, 20))

nv <- nrow(variables)
```
There are a total of `r nv` variable combinations that include our category of interest.
Note that `select_varid()` also allows selections by including other `ccmugs` to be more specific.
Let's inspect the variables in a more human-readable format:

```{r example-variablestext}
variables_text <- get_variables(
  ids = variables$variableId,
  rms = rem)
knitr::kable(variables_text[order(variables_text$classification, variables_text$measure), ])

```

Next, we filter for the desired measure `Net emissions/removals` and gas `Aggregate GHGs`:

```{r examples-variables-select}

var_ids <- variables_text[variables_text$measure=="Net emissions/removals" &
                 variables_text$gas=="Aggregate GHGs", ]
knitr::kable(var_ids)

```

All relevant information (e.g., units, type of gas) are provided here.
In addition, unit conversion factors can be identified by looking up relevant source and target units in `rem$units$units` and cross-referencing them in `rem$units$annexOne`.


#### **4. Query for download**

Queries using the Flexible Query API are done with the `.req` object in `rem`:

```{r example-query, eval = FALSE}
result <- flex_query(
  rem,
  variable_ids = var_ids$variable_id,
  party_ids = parties_id$id,
  year_ids = years$id,
  pretty = TRUE)

```


## Meta Overview

### Parties

Parties are stored in `rem$parties`

```{r parties, message=FALSE}

library(dplyr)
# up to 5th entry on parties
knitr::kable(rem$parties %>% 
               dplyr::group_by(name_code) %>% 
               dplyr::slice(1:3))

```

### Years


Years are stored in `rem$years` for respective 

```{r years, message=FALSE}

knitr::kable(names(rem$years))
knitr::kable(rem$years$annexOne)

```


## Variable Overview (*ccmug*)

- select directly from variable table
- or choose one `ccmgu` and find corresponding variables

```{r downloading}
# set up API object


```

Sections below give a brief overview on `ccmgu`'s.

#### Categories

Categories are UNFCCC CRF-Categories or categories under the Compilation and Accounting Data.
The tables for each category have an `id` and a `name` column.
Where nested sub-categories exist, additional columns using the name `level_x` (`x` = nesting depth) show the category hierarchy.


```{r categories}

knitr::kable(names(rem$categories))
# up to 5th entry on parties
knitr::kable(dplyr::slice(rem$categories$annexOne, 1:10))
knitr::kable(dplyr::slice(rem$categories$annexOne, 88:94))

```

#### Classification

Classification differ based on respective sectors and/or CRF-Categories:

- Energy: 
+ Fuel Types
- Agriculture:
+ Livestock types
- LULUCF/LUCF:
+ LULUCF/LUCF activities

```{r classification}

knitr::kable(head(rem$classification$annexOne, 10))

```

#### Measures

Measures include: 


```{r measures}


knitr::kable(head(rem$measures$annexOne, 10))

```


#### Gasses

Gasses include: 


```{r Gasses}


knitr::kable(head(rem$gas$annexOne, 10))

```


#### Units

Units include: 


```{r Units}


knitr::kable(head(rem$units$units, 10))

```

