---
title: "Interacting with the UNFCCC API using remis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interacting with the UNFCCC API using remis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`remis` uses the UNFCCC Data Interface Flexible Query API to download Inventory Data.
At a later stage, functions for downloading Compilation and Accounting Data (`cad` and `cadCP2`) will be included.  
Data downloads via queries require a variable for a specific party and year, which are referred to via unique `id`'s.
While party and year `id`'s are linked to respective data in a fairly straight-forward manner (e.g., single country or year), variable ids are unique descriptors for a combination of data sources consisting of:

- category
- classification
- measurement
- gas
- unit

The above are referred to as *ccmgu*'s.

Initializing `remis` with `rem_init` provides a list with several objects to inspect available data and query for it.
These include `id` overviews for parties, years, *ccmgu*'s, as well as respective 
variable tables and the query object for requests `.req`.


```{r setup, meassage = FALSE}
library(remis)

rem <- rem_init()

names(rem)
```

Note that only calls using the Flexible Query API are currently implemented. 
These allow interacting with data sources from:

```{r results='asis', echo=FALSE}

cat(paste0('- ', names(rem$categories[1:3]), collapse = '\n'))

```

Queries for data within `r paste(names(rem$categories[4:5]), collapse = ' and ')` will be available in future versions.


Further details on the objects within `rem` are:

```{r results='asis', echo=FALSE}
cat(
  paste0('- **',
         names(rem),
         '** - class: ',
         unlist(sapply(rem, function(x) paste(class(x), collapse = ', '))),
         collapse = '\n')
  )

```



## Building up Flexible Queries

Details on the Flexible Query API can be found in the official documentation of the [Data Interface for UNFCCC](https://unfccc.int/process-and-meetings/transparency-and-reporting/greenhouse-gas-data/data-interface-help#Flexible-Queries-).    

Define whether annexOne or nonAnnexOne data shall be downloaded, and then:

1. Select party
2. Select year
3. Choose variable based on `ccmgu`'s

Note, that objects in `rem` (excluding `parties`) provide all data sources (where applicable) reflected in the list items `r names(rem$categories)`.

### Parties

Parties are stored in `rem$parties`

```{r parties, message=FALSE}

library(dplyr)
# up to 5th entry on parties
knitr::kable(rem$parties %>% 
               dplyr::group_by(name_code) %>% 
               dplyr::slice(1:3))

```

### Years


Years are stored in `rem$years` for respective 

```{r years, message=FALSE}

knitr::kable(names(rem$years))
knitr::kable(rem$years$annexOne)

```


### Variables

- select directly from variable table
- or choose one `ccmgu` and find corresponding variables

```{r downloading}
# set up API object


```

Sections below give a brief overview on `ccmgu`'s.

#### Categories

Categories are UNFCCC CRF-Categories or categories under the Compilation and Accounting Data.
The tables for each category have an `id` and a `name` column.
Where nested sub-categories exist, additional columns using the name `level_x` (`x` = nesting depth) show the category hierarchy.


```{r categories}

knitr::kable(names(rem$categories))
# up to 5th entry on parties
knitr::kable(dplyr::slice(rem$categories$annexOne, 1:10))
knitr::kable(dplyr::slice(rem$categories$annexOne, 88:94))

```
  
#### Classification

Classification differ based on respective sectors and/or CRF-Categories:

- Energy: 
  + Fuel Types
- Agriculture:
  + Livestock types
- LULUCF/LUCF:
  + LULUCF/LUCF activities
  
```{r classification}

knitr::kable(head(rem$classification$annexOne, 10))

```

#### Measures

Measures include: 


```{r measures}


knitr::kable(head(rem$measures$annexOne, 10))

```


#### Gasses

Gasses include: 


```{r Gasses}


knitr::kable(head(rem$gas$annexOne, 10))

```


#### Units

Units include: 


```{r Units}


knitr::kable(head(rem$units$units, 10))

```

